
##Coursera Statistical Inference - Simulation-Part 2
####Author: "Joanna Teo"


## Overview
Report analyses the ToothGrowth data as per the instructions provided 
In Coursera Statistical Inference, Basic inferential data analysis.

```{r load.other, echo=FALSE}
library(ggplot2)
library(plyr)
```

## Load the data
```{r load, echo=TRUE}
library(datasets)
data(ToothGrowth)
```
## Analysis
### 1. Basic exploratory data analysis
Have a quick glance what this data is like
```{r str, echo=TRUE}
str(ToothGrowth)
```
It's not immediately self explanatory, reading the help using `help(ToothGrowth)` we gather the following information

We have 3 columns in `ToothGrowth dataset` 

|   Name   |   Type    |    Values           | 
|----------|-----------|---------------------|
|   `len`  |   numeric |  Tooth length       |
|   `supp` |   factor  |  **VC** or **OJ**   |
|   `dose` |   numeric | Dose in milligrams |

And it tracks the response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid, OJ or VC respectively).
```  
  
### 2. Basic summary 
  
Observe R summary result
```{r summary, echo=TRUE}  
summary(ToothGrowth)
````

Inspect the tests distribution by supplement and dose
```{r test.distribution, echo=TRUE}
table(ToothGrowth$dose, ToothGrowth$supp)
```
10 tests for each pair.

```{r plot, echo=TRUE}
agg <- aggregate(len ~ dose + supp, ToothGrowth, mean)
ggplot(agg, aes(x=dose, y=len, colour = supp)) +
    geom_line(size=2, alpha=.5) + geom_point(size=5, alpha=.3) +
    xlab("Dose(miligrams)") + ylab("Avg. Tooth length") +
    guides(colour=guide_legend(title="Supplement type")) +
    scale_color_manual(values = c("darkblue", "#E69F00"))
```  
  
There seems to be a correlation between the dose and tooth growth, where 
Orange Juice is more effective for lower doses, where the 2 milligram 
seems to be the maximum effect point.

### 3. Comparisom of tooth growth by supp and dose  
To verify that that our conclusions based on the sample diagram, we need 
identify the confidence interval for each of the supplement/dose.
```{r by.dose, echo=TRUE}
ddply(ToothGrowth, dose ~ supp,function(x) 
      c(mean=mean(x$len), sd=sd(x$len),
        conf.int=t.test(x$len)$conf.int))
```  
We observe that in 95% confidence interval the `Ascorbic Acid(VC)` intervals are pairwise disjoint so we can claim with high level of confidence that the length means are distinct, moreover there is a clear growth correlation between dose & length means.

By now we can also immediately identify with high level of confidence 
that For 0.5 and 1 milligrams Orange Juice have has greater impact on 
tooth growth (On the merit that for those 2 doses there confidence
interval are pairwise disjoint).

For Orange Juice(OJ) supplement type, however, there is an overlap for dose 1 and 2 milligrams, 
and we are forced to look deeper.
```{r by.dose.hypothesis.test, echo=TRUE}
t.test(len ~ dose, paired=FALSE, var.equal=TRUE,
       data=subset(ToothGrowth, dose %in% c(1.0,2.0) & supp == 'OJ'))
```

```{r by.dose.hypotheis.test.hidden, echo=FALSE}
hp <- t.test(len ~ dose, paired=FALSE, var.equal=TRUE,
       data=subset(ToothGrowth, dose %in% c(1.0,2.0) & supp == 'OJ'))
t <- hp$statistic
p <- hp$parameter
q <- qt(.025, p)
```
The t value ```r t``` being less than **qt(.025, ```r p```)** == ```r q```
allows us to assert that the mean length for 2 milligrams as greater than the for the 1 milligram dose.

In the 2.0 milligram dose there is an overlap between `Orange Juice (OJ)` and `Ascorbic Acid (VC)` let's dig deeper 
```{r by.supp, echo=TRUE}
t.test(len ~ supp, paired=FALSE, var.equal=FALSE,
       data=subset(ToothGrowth, dose == 2.0))
```
The confidence Interval includes 0 and hence difference between the supplements types vis-a-vis mean lengths is insignificant.

## Conclusion
We have conducted an analysis of the ToothGrowth data based on the 
assumptions that guinea pigs were randomly chosen and hence the 
population variance the same. We are also assuming that the samples
are independent so the unpaired testing was employed.

Our analysis has shown with high confidence that the there is a 
correlation between the supplement type used and teeth growth in guinea pigs, 
when for small doses of 0.5 and 1 milligrams, Orange Juice, clearly has
an advantage.
