---
##title: "Coursera Statistical Inference - Simulation"
###author: "Joanna Teo"
###date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
   : united
    toc: yes highlight: pygments
    keep_md: yes
    theme

```{r required.packages, echo=FALSE}
require('knitr', quietly=TRUE)
```

## Overview
We will perform an exponential distribution simulation and 
compare it to the CLT(Centeral Limit Theorem). The excercise we are 
undertaking is already restricted to specific values that we are 
documented in the code itself.

## Simulations
As instructed we are going to perform 1000 simulations of 40 samples each, where `lambda = 0.2`

```{r simulations}
set.seed(1234)      # Just a seed 
lambda <- 0.2       # Given by the exercise
sample.size <- 40   # Given by the exercise
sims <- 1000        # Suggested by the exercise
d.mean <- 1/lambda  # Distribution mean. Inherent property
d.sd   <- 1/lambda  # Disribution standard deviation. Inheret property

exp.sim <- matrix(rexp(sims*sample.size, rate=lambda), sims, sample.size)
means <- rowMeans(exp.sim)
```  
  
## Analysis
```{r diagram.function, echo=FALSE}
plot.hist.scale <- function(subject, ms, ss) {
  hist(ms, breaks=25, probability=TRUE, main=subject, xlab="", col="darkgreen")
  lines(density(ms), col="red", lwd=3) 
  curve(dnorm(x, d.mean, d.sd/sqrt(ss)), min(ms), max(ms), col="purple", add=T, lty=2, lwd=3)

  # add legend
  legend('topright', c("simulation", "theoretical"), lty=c(1,2), col=c("red", "purple"))
}  
  
plot.hist <- function(subject, ms, ss) {
  hist(ms, breaks=25, prob=TRUE,
       main=subject, xlab="", col="darkgreen")
  lines(density(ms), col="red", lwd=3) # density of the averages of samples

  abline(v=d.mean, col="darkblue", lwd=3) # Theoretical distribution center

  # theoretical density of the averages of samples
  curve(dnorm(x, d.mean, d.sd/sqrt(ss)), min(ms), max(ms), col="purple", add=T, lty=2, lwd=3)
  # add legend
  legend('topright', c("simulation", "theoretical"), lty=c(1,2), col=c("red", "purple"))
}
```

```{r diagram, echo=FALSE}
plot.hist("Sample averages' of Exp. distribution. with rate = 0.2", means, sample.size)
``` 

### Sample mean vs. theoretical mean
```{r mean, echo=TRUE}
sample.mean <- mean(means)
theoretical.mean <- 1/lambda
```  
The deviation of the sample mean ```r sample.mean``` from the theoretical
mean ```r theoretical.mean``` is approximately 
```r round(abs(sample.mean - theoretical.mean)/theoretical.mean, 3) * 100```% 
      
### 2. Sample variance vs. Theoretical variance  
  
```{r variance, echo=TRUE}
sample.sd <- sd(means)
theoretical.sd <- (d.sd)/sqrt(sample.size)

sample.var <- var(means)
theoretical.var <- 1 / (lambda ** 2 * 40)
```  
The sample variance ```r sample.var``` is different from the theoretical variance ```r theoretical.var``` by 
```r round(abs(sample.var - theoretical.var)/ theoretical.var, 3)*100```%

and the sample standard deviation ```r sample.sd``` diffrantiates from the 
theoretical standard deviation ```r theoretical.sd``` by 
```r round(abs(sample.sd - theoretical.sd)/ theoretical.sd, 3)*100```%
  
### 3. Is the distribution normal?  

```{r samples, echo=FALSE, fig.width=12, fig.height=8}
mult = 20
par(mfrow=c(1,2))
means2 <- rowMeans(matrix(rexp(sims*sample.size*mult, rate=lambda), 
                 sims, sample.size * mult))
plot.hist("Sample averages' (40 samples)", means, sample.size)
plot.hist("Sample averages (800 samples)", means2, sample.size * mult)
```
**Notice** how the the averages of 800 samples from the exponential distrbution converges to the mean vis-a-vis the the averages of 40 samples, it is also visibly
closer to the normal distribution. 

#### Q-Q Plot 

```{r normal, echo=FALSE, fig.height=3, fig.width=4}    
qqnorm(means, col="darkgreen")
qqline(means, col="purple")
```  

A Q-Q plot indicates as well that the 40 exponentials are quite close to the normal distribution.
